---
title: "Main_Analysis"
authors: "Aarav Gosalia, Aleric Govender, Jordi Capdevila Maso"
date: "2025-03-19"
output: pdf_document
---

## EDA

```{r}
# Load necessary libraries
library(readr)
library(tidyverse)

# 1. Load and Clean Data
csv_file <- "reddit_engagement_data.csv"
df <- read_csv(csv_file, show_col_types = FALSE)
df_cleaned <- df %>% 
  drop_na()

df_cleaned <- df_cleaned %>%
  mutate(
    `Subreddit Name` = as.factor(`Subreddit Name`),
    `Contains Question` = as.factor(`Contains Question`),
    `Contains Emoji` = as.factor(`Contains Emoji`),
    `Contains Profanity` = as.factor(`Contains Profanity`),
    `Is Early Comment` = as.factor(`Is Early Comment`),
    `Comment Hour` = as.factor(`Comment Hour`),
    `Comment Day` = as.factor(`Comment Day`)
  )

names(df_cleaned) <- gsub(" ", "_", names(df_cleaned))
names(df_cleaned) <- gsub("[()]", "", names(df_cleaned))

# Optional: write cleaned data to file
#write_csv(df_cleaned, "reddit_cleaned.csv")

# 2. Basic Summary Statistics
summary(df_cleaned)

# 3. Full Distribution of Comment Score
ggplot(df_cleaned, aes(x = Comment_Score)) +
  geom_histogram(bins = 50, fill = "lightblue", color = "white") +
  labs(
    title = "Histogram of Comment Scores",
    x = "Comment Score",
    y = "Count"
  ) +
  theme_minimal()

# 4. Zoomed-In Distribution (-10 to 50)
df_filtered <- df_cleaned %>% filter(Comment_Score >= -10 & Comment_Score <= 50)
ggplot(df_filtered, aes(x = (Comment_Score))) +
  geom_histogram(fill = "darkblue", bins = 60, alpha = 0.7) +
  theme_minimal() +
  labs(title = "Truncated Distribution of Comment Scores",
       x = "Comment Score",
       y = "Frequency") +
  theme(plot.title = element_text(hjust = 0.5))

```
## Inverse Hyperbolic Sine (IHS) Transformation Approach

```{r}
# Distribution of Transformed Comment_Score using asinh() 
ggplot(df_cleaned, aes(x = asinh(Comment_Score))) +
  geom_histogram(fill = "darkviolet", bins = 60, alpha = 0.7) +
  theme_minimal() +
  labs(title = "Distribution of IHS Transformed Comment Scores",
       x = "asinh(Comment Score)",
       y = "Frequency") +
  theme(plot.title = element_text(hjust = 0.5))
```

## Shifted Score Approaches

```{r}
shift_df <- df_cleaned
lowest_score <- min(shift_df$Comment_Score)
shift_df$Comment_Score_shift <- shift_df$Comment_Score + abs(lowest_score) + 1
```


### Log-Linear Multiple Linear Regression

```{r}
library(MASS)

shift_df$log_score <- log(shift_df$Comment_Score_shift)

#Force Sentiment_Score to be part of the model
model_null_mlr <- lm(log_score ~ Sentiment_Score, data = shift_df)

#model_null_mlr <- lm(log_score ~ 1, data = shift_df)

model_full_mlr <- lm(log_score ~ Sentiment_Score + Text_Length + Word_Count + 
                   Contains_Question + Contains_Emoji + Contains_Profanity + 
                   Comment_Age_hours + Comment_Hour + Comment_Day + 
                   Is_Early_Comment + Parent_Score + User_Karma + 
                   Account_Age_days + Subreddit_Name, data = shift_df)

step_model_mlr <- stepAIC(model_null_mlr, 
                          scope = list(lower = model_null_mlr, upper = model_full_mlr), 
                          direction = "forward", 
                          trace = TRUE)

summary(step_model_mlr)

par(mfrow=c(2,2))
plot(step_model_mlr)
par(mfrow=c(1,1))
```
### Gamma Regression with Log Link

```{r}
#Force Sentiment_Score to be part of the model
model_null_gamma <- glm(Comment_Score_shift ~ Sentiment_Score, family = Gamma(link = "log"), data = shift_df)

#model_null_gamma <- glm(Comment_Score_shift ~ 1, family = Gamma(link = "log"), data = shift_df)

model_full_gamma <- glm(Comment_Score_shift ~ Sentiment_Score + Text_Length + Word_Count + 
                   Contains_Question + Contains_Emoji + Contains_Profanity + 
                   Comment_Age_hours + Comment_Hour + Comment_Day + 
                   Is_Early_Comment + Parent_Score + User_Karma + 
                   Account_Age_days + Subreddit_Name, family = Gamma(link = "log"), data = shift_df)

step_model_gamma <- stepAIC(model_null_gamma, 
                            scope = list(lower = model_null_gamma, upper = model_full_gamma), 
                            direction = "forward", 
                            trace = TRUE)

summary(step_model_gamma)

par(mfrow=c(2,2))
plot(step_model_gamma)
par(mfrow=c(1,1))
```
### Generalized Additive Model (GAM)

```{r}
library(mgcv)

gam_model <- gam(Comment_Score_shift ~ 
                   s(User_Karma) +
                   Comment_Age_hours + 
                   s(Word_Count) + 
                   s(Account_Age_days) + 
                   s(Text_Length) +
                   s(Parent_Score) +
                   s(Sentiment_Score) + 
                   Is_Early_Comment + 
                   Subreddit_Name + 
                   Comment_Day + 
                   Comment_Hour + 
                   Contains_Question + 
                   Contains_Emoji + 
                   Contains_Profanity, data = shift_df)

summary(gam_model)

plot(gam_model, pages = 1, shade = TRUE)

par(mfrow=c(2,2))
gam.check(gam_model)
par(mfrow=c(1,1))
```

##Fitting Other Shifting Based Models (Not used in paper)

### Poisson Model

```{r}
full.model <- glm(Comment_Score_shift ~ Sentiment_Score + Text_Length + Word_Count + 
                   Contains_Question + Contains_Emoji + Contains_Profanity + 
                   Comment_Age_hours + Comment_Hour + Comment_Day + 
                   Is_Early_Comment + Parent_Score + User_Karma + 
                   Account_Age_days + Subreddit_Name, data = shift_df, family = poisson)

summary(full.model)

par(mfrow=c(2,2))
plot(full.model)
par(mfrow=c(1,1))
```
### Negative Binomial Model

```{r}
nb.model <- glm.nb(Comment_Score_shift ~ Sentiment_Score + Text_Length + Word_Count + 
                   Contains_Question + Contains_Emoji + Contains_Profanity + 
                   Comment_Age_hours + Comment_Hour + Comment_Day + 
                   Is_Early_Comment + Parent_Score + User_Karma + 
                   Account_Age_days + Subreddit_Name, data = shift_df)
summary(nb.model)

par(mfrow=c(2,2))
plot(nb.model)
par(mfrow=c(1,1))
```
## Extra : Fitting IHS models (Not used in report)

### Multiple Linear Regression (MLR)

```{r}
library(MASS)

# Create the IHS-transformed response variable (using asinh)
df_cleaned <- df_cleaned %>% 
  mutate(ihs_score = asinh(Comment_Score))

model_null <- lm(ihs_score ~ 1, data = df_cleaned)

model_full <- lm(ihs_score ~ Sentiment_Score + Text_Length + Word_Count + 
                   Contains_Question + Contains_Emoji + Contains_Profanity + 
                   Comment_Age_hours + Comment_Hour + Comment_Day + 
                   Is_Early_Comment + Parent_Score + User_Karma + 
                   Account_Age_days + Subreddit_Name , data = df_cleaned)

step_model <- stepAIC(model_null, 
                      scope = list(lower = model_null, upper = model_full), 
                      direction = "forward", 
                      trace = TRUE)

summary(step_model)

par(mfrow = c(2, 2))
plot(step_model)
par(mfrow = c(1, 1))
```
### GAM (2)

```{r}
gam_model <- gam(ihs_score ~ 
                  s(Comment_Age_hours) + 
                  s(User_Karma) + 
                  s(Word_Count) + 
                  s(Account_Age_days) + 
                  s(Sentiment_Score) + 
                  s(Text_Length) +
                  s(Parent_Score) +
                  Is_Early_Comment + 
                  Subreddit_Name + 
                  Comment_Day + 
                  Comment_Hour + 
                  Contains_Question + 
                  Contains_Emoji + 
                  Contains_Profanity,
                data = df_cleaned, 
                select = TRUE)

summary(gam_model)

plot(gam_model, pages = 1, shade = TRUE)

par(mfrow=c(2,2))
gam.check(gam_model)
par(mfrow=c(1,1))
```
